### 2019-01-21 writing/thesis/chapters/methods.tex
Final (?) changes.

% TODO: Extract to settings
\chapter{An ADER-DG scheme for the Navier-Stokes Equations}\label{chap:methods}
This chapter describes the numerical and physical background that is needed to simulate reacting fluids with high accuracy.
We start with a description of the arbitrary high order derivatives (\ader{}) discontinuous Galerkin (\dg{}) method, with a focus on equations that contain both advective and diffusive terms.
We then segue into a short description of the \muscl{} scheme, which is a second order finite volume method.

@@

The hyperbolic flux is given by
\begin{equation}
  \label{eq:hyper-flux}
  \hyperFlux(\Q) = \hyperFluxDef,
\end{equation}
which are the Euler equations coupled with an advection equation.
In this equation, $\bm{a} \otimes \bm{b} = \bm{a} \bm{b}^\intercal$ is the Kronecker or outer product of two vectors $\bm{a}$ and $\bm{b}$.
The pressure $\pressure$ is given by the equation of state of an ideal reacting gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - \frac{1}{2} \left(\Qv \cdot \Qj \right)  - q_0 \QZ - gz \right).
\end{equation}
The term $q_0 \QZ$ is the chemical energy with a heat release $q_0$~\cite{helzel2000modified}, the term $gz$ is the geopotential height~\cite{giraldo2008study}.
We set $g$ to 9.81.
The pressure is related to the temperature $T$ by the thermal equation of state
\begin{equation}
  \label{eq:temperature}
  \pressure = \Qrho R T,
\end{equation}
where $R$ is the specific gas constant of a fluid.

The viscous flux is
\begin{equation}
  \label{eq:visc-flux}
  \viscFlux(\Q, \gradQ) = \viscFluxDef,
\end{equation}
where $\stressT$ and $(\kappa \nabla T)$ denote the stress tensor and heat flux respectively.
The viscous effects of the fluid are modeled by the stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) =
  \mu
  \left(
  \left(\nicefrac{2}{3} \divergence{\Qv} \right) -
  \left( \gradient{\Qv} + \gradient{\Qv}^\intercal \right)
  \right).
\end{equation}

We further introduce the ratio of specific heats $\gamma$ and the heat fraction for constant volume $c_v$ and constant volume $c_p$.
These constants are all fluid dependent and relate to each other~\cite{dumbser2010arbitrary} and the gas constant by
\begin{align}
  \label{eq:fluid-constants}
  \begin{split}
  c_v &= \frac{1}{\gamma - 1} R \\
  c_p &= \frac{\gamma}{\gamma - 1} R\\
  R &= c_p - c_v\\
  \gamma &= \frac{c_p}{c_v}.
  \end{split}
\end{align}
We can compute the heat conduction coefficient
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R = \frac{\mu c_p}{\Pr},
\end{equation}
where the Prandtl number $\Pr$ depends on the fluid.
The temperature gradient $\gradient{T}$ in terms of the conservative variables can be found in \cref{chap:appendix-pde}.

@@

We also support two different kind of optional source terms.
The first source term is gravity, given by
\begin{equation}\label{eq:source-gravity}
  \source[\Qj] = - g \bm{e_z} \cdot \Qj,
\end{equation}
where $\bm{e_z}$ is the unit vector pointing in $z$-direction~\cite{giraldo2008study}.

@@

The chemical reactive source term is
\begin{equation}\label{eq:source-reaction}
  \source[\QZ] = - K(T) \QZ.
\end{equation}
We use the simple reaction model
\begin{equation}\label{eq:reaction-model}
\begin{cases}
  - \frac{1}{\reactionTimescale} & T > \reactionTemperature\\
  0 & \text{otherwise}
\end{cases}  
\end{equation}
where $\reactionTimescale$ is the timescale of the reaction and $\reactionTemperature$ is the activation temperature~\cite{hidalgo2011ader,helzel2000modified}.

@@

The \aderdg{}-scheme is a predictor-corrector method.
We first compute a local time evolution of each cell in a predictor step and then add the influence of the neighbors in the corrector step.
\sidetitle{Reference Cell}
For reasons of computational efficiency, we describe the scheme in terms of a reference cell.
We use the quad cell $\hat{\cell} = (0, 1)^\dimensions$ as space reference cell.
The mapping, which takes a point on the reference cell with coordinates $\hat{\bm{x}}$ and and returns its coordinates $\bm{x}$ in a grid cell $\cell$ is
\begin{equation}\label{eq:space-mapping}
  \bm{x} = \mapping (\hat{\bm{x}}) =
  \cellCenter(\cell) +
\begin{pmatrix}
\Delta x && \\
&\Delta y & \\
&&\Delta z&
\end{pmatrix}
  \begin{pmatrix}
    \hat{x} - 0.5\\
    \hat{y} - 0.5\\
    \hat{z} - 0.5
  \end{pmatrix}.
\end{equation}
The symbol $\mapping^{-1}$ denotes the inverse mapping.
It can be used to define a function $f(\bm{x})$ acting on a point $\hat{\bm{x}}$ of a regular cell in terms of a function $\hat{f}(\bm{x})$ that acts on the corresponding point of the reference cell by the relation
\begin{equation}
  \label{eq:function-ref-cell}
  f(\bm{x}) = f\left( \mapping^{-1}(\bm{x}) \right) = \hat{f}(\hat{\bm{x}}).
\end{equation}

@@

The Jacobian determinant of $\mapping$ is simply the volume $V$ of the cell
\begin{equation}
  \label{eq:determinant-mapping}
  \volume = \operatorname{det}(\operatorname{Jacobian}\mapping) = \Delta x \, \Delta y \,\Delta z,
\end{equation}
the Jacobian determinant of $\timeMapping$ is $\Delta t$.
\todo{Similarly -> as well as?}
This is useful for evaluating derivatives by the chain rule and, similarly, integrals by substitution.
For example, we can evaluate a volume integral over a cell by
\begin{equation}
  \label{eq:integration-by-substitution}
  \intdcell{
f(\bm{x})
  }
  =
\volume \intdrefcell{
    \hat{f}(\hat{\bm{x}})
  }.
\end{equation}
\sidetitle{Basis Functions}
The next step is the definition of basis functions for our cells.
We use Lagrange polynomials that are defined in the one-dimensional case for the reference interval $(0,1)$ by
\begin{align}
  \label{eq:lagrange-basis}
    \lagrangeRef(x) = \prod_{\substack{0 \leq j \leq N \\ i \neq j}} \frac{x - {\quadNode[j]}}{x_i - \quadNode[j]}.
\end{align}
We can then approximate functions as a sum over the support points $\quadNode[j]$
\begin{equation}
  \label{eq:langrange-expansion-1d}
  f(x) = \sum_{i = 0}^N f(x_i) \lagrangeRef[i](x) = f(x_i) \lagrangeRef[i](x),
\end{equation}
where we used the Einstein summation convention which implies summation over repeated free indices.
This convention is used from here on.
The support points $\quadNode[i]$ are chosen such that they coincide with the nodes of the Gauss-Legendre quadrature of order $N+1$ which we can use to evaluate integrals
\begin{equation}
  \label{eq:quadrature-1d}
  \int_0^1 f(\hat{x}) \dd{\hat{x}} = \sum_{i=0}^{N} \quadWeight[i] f(\quadNode[i]),
\end{equation}
where $\quadWeight[i]$ is the quadrature weight corresponding to the node $\quadNode[i]$.

The basis functions are similarly defined in $\dimensions$-dimensions by considering nodes that are the tensor-products of $d$ one-dimensional Gauss-Legendre quadrature nodes.
They are indexed by a multi-index $n = (n_1, \ldots, n_\dimensions)$
\begin{equation}
  \label{eq:basis-space}
  \sbasisRef[][n] (\bm{x}) = \prod_{i=0}^{\dimensions} \lagrangeRef[n_i] (x_i)
\end{equation}
that is linearized such that $n = (0, \ldots, \sbasisSize)$.
A $d$-dimensional function is then approximated by
\begin{equation}
  \label{eq:lagrange-expansion}
  f(\bm{x}) = \sum_{i = 0}^{\sbasisSize - 1} f(\bm{x}_i) \lagrangeRef[i](\bm{x}) = f(\bm{x}_i) \lagrangeRef[i](\bm{x}).
\end{equation}
Similarly to the one-dimensional case, we can evaluate integrals by
\begin{align}
  \label{eq:quadrature}
  \underbrace{\int_0^1 \cdots \int_0^1}_{\dimensions \text{ times}} f(\hat{\bm{x}}) \dd{\hat{\bm{x}}} &=
 \sum_{n=0}^{\sbasisSize-1} \quadWeight[n] f(\quadNodeNd[n]) \\
  \intertext{with}
  \quadWeight[n] &= \prod_{i}^{d} \quadWeight[n_i],
\end{align}
where the quadrature weights $\quadWeight[n]$ are the product of all relevant one-dimensional quadrature weights, using the aforementioned (linearized) multi-index $n$.

@@

We define the basis functions over a space-cell with an index $n = 0, \ldots, \sbasisSize - 1$
\begin{align}\label{eq:space-basis}
  \begin{split}
  \sbasis[][n] (\bm{x}) &= \sbasisRef[][n] \left( \mapping^{-1}(\bm{x}) \right).
  \end{split}
\end{align}
In addition to the space cells, we also introduce space-time cells $\cell \times [\curTime, \nextTime]$ with corresponding reference cell $(0,1)^\dimensions \times  [0,1]$.
A basis for these cells is given by
\begin{align}\label{eq:space-time-basis}
  \begin{split}
  \stbasisRef[][l,n](\hat{\bm{x}}, \hat{t}) &= \lagrange[l](t) \sbasis[][n] (x)\\
  \stbasis[][l,n] (\bm{x}, t) &= \stbasisRef \left( \mapping^{-1}(\bm{x}), \timeMapping^{-1}(t) \right)
  \end{split}
\end{align}
where the first index $l = 0, \ldots, N$ runs over the time.
Outside of elements, we define the basis functions to be equal to zero.

@@

\sidetitle{Predictor}
We are now ready to derive the predictor.
We first multiply the conservation law by a test function of the same function space as the basis, integrate over a space-time-element $\cell$ and replace all terms with their discrete approximations.
In particular, we replace the vector of conservative variables $\Q$ with the space-time predictor $\stpredictor[\cell]$.
\begin{align}\label{eq:weak-pde-space-time}
\begin{split}
\intdt{\intdcell{
    \sttestfunction[\cell][v,l,n](\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
&+
\intdt{\intdcell{
    \sttestfunction[\cell][v,l,n](\bm{x}, t)
    \left(
      \divergence{\stflux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
\\&=
\intdt{\intdcell{
  \sttestfunction[\cell][v,l,n] \stsource(\bm{x}, t, \stpredictor)
}}.
\end{split}
\end{align}
%where the indices $v = (0, \ldots, \NVar-1)$, $n = (0, \ldots, \sbasisSize - 1)$ and $l = (0, \ldots, N)$ run over the variables, space and time.

Henceforth, we drop the cell index for all quantities and the spatial argument for the basis and test functions.
We integrate the first term of \cref{eq:weak-pde-space-time} by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the known discrete solution at time $t$.
This corresponds to upwinding in time~\cite{dumbser2008unified}.
Note that this neglects the interaction with neighboring cells.
The cell-local scheme is then
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \sttestfunction[][v,l,n] (\nextTime) \stpredictor[i+1](\nextTime)
}
&-
\intdt{\intdcell{
    \pdv{}{t} \sttestfunction[][v,l,n] (t) \stpredictor[i+1](t)
}}
= 
\intdcell{
  \sttestfunction[][v,l,n] (\curTime) \ssol (\curTime)
}
\\&+
\intdt{\intdcell{
    \sttestfunction[][v,l,n] (t) \divergence{\stflux(\stpredictor[i], \gradient{\stpredictor[i]})}
}}
\\&+
\intdt{\intdcell{
    \sttestfunction[][v,l,n] (t) \stsource(\stpredictor[i])
}}.
\end{split}
\end{align}
Here, we introduced the iteration counter $i$ as a subscript for the space-time predictor.
Using the expansions in space and time~\pcref{eq:cell-approx-space,eq:cell-approx-space-time} results in a local systems of equations that can be solved by a fixed-point iteration scheme.
We set the initial value for the space-time predictor as
\begin{equation}
  \label{eq:initial-guess}
\quad \stpredictorCoeff[0][v,l,n] = \ssolCoeff[][v,n]
\end{equation}
which corresponds to a stationary initial guess~\cite{dumbser2008unified}.
For an overview of possible initial guesses, see~\cite{dumbser2018efficient}.

@@

\sidetitle{Corrector}
The final step of the \aderdg{}-scheme is the corrector.
We multiply the system \cref{eq:conservation-law-gradient} by a space test function and integrate over the space-time control volume.
This results, yet again, in the weak formulation of the \pde{}
  \label{eq:weak-pde}
\intdt{\intdcell{
\stestfunction[][v,n] \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \stestfunction[][v,n] \left( \div{F(\Q, \gradQ } \right)
}}
=
\intdt{\intdcell{
    \stestfunction[][v,n] S(\Q, \bm{x}, t)
}}.
We now insert our discrete approximations and apply the space-basis expansion~\pcref{eq:cell-approx-space} to the discrete solution.
The first term can be integrated by parts in time.
This has the convenient result that only the boundary terms remain because the basis is constant in time.
We then integrate the flux divergence by parts in space and insert the Riemann solver to connect with the neighboring elements.
Finally, we need to solve
\newcommand{\massMatrixDef}{\intdcell{
  \stestfunction[][v,n] \sbasis[][v,n']
}}%
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\ssolCoeff[\nextTime][v,n'] - \ssolCoeff[\curTime][v,n'])
&+
\left(\intdt{\intdcellb{
      \stestfunction[][v,n] \Riemann(\stpredictor[-], \gradient{\stpredictor[-]}; \stpredictor[+], \gradient{\stpredictor[+]}) \cdot \normal
}}\right)
\\&-
\left(\intdt{\intdcell{
    \gradient{\stestfunction[][v,n]} \cdot  \stflux(\spredictor, \gradient{\spredictor})
}}\right)
\\&=
\left(\intdt{\intdcell{
      \stestfunction[][v,n] \stsource(\spredictor)
}}\right)
for the coefficients $\ssolCoeff[\nextTime][v,n]$ of the new solution.
Here the superscripts $\pm$ denote values that are extrapolated to the boundaries.
Again, the computational effort can be vastly reduced by collecting the integrals into matrices and precomputing as much as possible.
Using the orthogonality of the basis functions~\pcref{eq:basis-orthogonality} on the first term of \cref{eq:corrector}, it becomes clear that the correction is a simple, explicit computation.
% The first term is the mass tensor with entries
% \newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
% \begin{equation}
%   \label{eq:mass-matrix}
%   \massMatrix[v n n'] = \massMatrixDef = \volume \underbrace{\quadWeight[n] \delta_{n n'}}_{\text{diagonal}}
% \end{equation}
% which is, due to the orthogonal basis \pcref{eq:basis-orthogonality}, diagonal for each variable $v$ and thus trivial to invert.
% We can then insert the definition of our basis function into our equation and arrive at
% \begin{align}
%   \massMatrix \left( \ssolCoeff[t+1] - \ssolCoeff[t] \right) &=
%   \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
%   \intertext{with}
% a_{v,n} &= \sfluxCoeff[][n'] \intdcell{
%   \sbasis[][v,n'] \gradient{\sbasis[][v,n]}(t)
% }
% \\
% \label{eq:corrector-surface}
% b_{v,n} &= \intdcellb{
%     \stestfunction[][v,n] \Riemann \left( \spredictor[-], \gradient{\spredictor[-]}; \spredictor[+], \gradient{\spredictor[+]} \right) \cdot \normal
% } \\
% s_{v,n} &=
% \ssourceCoeff[][v,n'] \intdcell{
%   \sbasis[][v,n] \sbasis[][v,n']
% }
% ,
% \end{align}
% where we evaluated the time-integrals by \cref{eq:integration-by-substitution}.
% All quantities are now time-averaged~\pcref{eq:time-average} and are represented in the space basis~\pcref{eq:cell-approx-space}.
Note that we evaluate the Riemann problem for the time-averaged space-predictor in our implementation.
This is justified by the linearity of our numerical flux~\cite{dumbser2008unified,dumbser2010arbitrary}.
% The surface integral \pcref{eq:corrector-surface} can be evaluated by summing over the contributions of the faces of each cell, which are computed by integrating the extrapolated space predictor with Gaussian quadrature.
% Finally, we can solve \cref{eq:corrector} for the new unknows $\ssolCoeff[k+1]$
% \begin{align}\label{eq:update-predictor}
% \begin{split}
%   %\bm{u}^{n+1}
%   \ssolCoeff[k+1][v]
%   &= \ssolCoeff[k][v] + \Delta t (\operatorname{update}) \\
%   \operatorname{update} &= \massMatrix[v]^{-1} \left( \bm{a}_v - \bm{b}_v + \bm{s}_v \right)
% \end{split}
% \end{align}
% which is clearly of the form of a one-step scheme.

@@

>>>>

% TODO: Extract to settings
\chapter{An ADER-DG Scheme for the Navier-Stokes Equations}\label{chap:methods}
This chapter describes the numerical and physical background that is needed to simulate reacting fluids with high accuracy.
We start with a description of the arbitrary high order derivatives (\ader{}) discontinuous Galerkin (\dg{}) method, with a focus on equations that contain both advective and diffusive terms.
We then segue into a short description of the \muscl{} scheme, which is a second order finite volume method.

@@

\sidetitle{Hyperbolic flux}
The hyperbolic flux is given by
\begin{equation}
  \label{eq:hyper-flux}
  \hyperFlux(\Q) = \hyperFluxDef,
\end{equation}
which are the Euler equations coupled with an advection equation.
In this equation, $\bm{a} \otimes \bm{b} = \bm{a} \bm{b}^\intercal$ is the Kronecker or outer product of two vectors $\bm{a}$ and $\bm{b}$.
The pressure $\pressure$ is given by the equation of state of an ideal reacting gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - \frac{1}{2} \left(\Qv \cdot \Qj \right)  - q_0 \QZ - gz \right).
\end{equation}
The term $q_0 \QZ$ is the chemical energy with a heat release $q_0$~\cite{helzel2000modified}, the term $gz$ is the geopotential height with the gravity of Earth $g$~\cite{giraldo2008study}.
The pressure is related to the temperature $T$ by the thermal equation of state
\begin{equation}
  \label{eq:temperature}
  \pressure = \Qrho R T,
\end{equation}
where $R$ is the specific gas constant of a fluid.

\sidetitle{Viscous flux}
The viscous flux is
\begin{equation}
  \label{eq:visc-flux}
  \viscFlux(\Q, \gradQ) = \viscFluxDef,
\end{equation}
where $\stressT$ and $(\kappa \nabla T)$ denote the stress tensor and heat flux respectively.
The viscous effects of the fluid are modeled by the stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) =
  \mu
  \left(
  \left(\nicefrac{2}{3} \divergence{\Qv} \right) -
  \left( \gradient{\Qv} + \gradient{\Qv}^\intercal \right)
  \right).
\end{equation}

We further introduce the ratio of specific heats $\gamma$ and the heat fraction for constant volume $c_v$ and constant volume $c_p$.
These constants are all fluid dependent and relate to each other~\cite{dumbser2010arbitrary} and the gas constant by
\begin{align}
  \label{eq:fluid-constants}
  \begin{split}
  c_v &= \frac{1}{\gamma - 1} R, \\
  c_p &= \frac{\gamma}{\gamma - 1} R,\\
  R &= c_p - c_v,\\
  \gamma &= \frac{c_p}{c_v}.
  \end{split}
\end{align}
We can compute the heat conduction coefficient
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R = \frac{\mu c_p}{\Pr},
\end{equation}
where the Prandtl number $\Pr$ depends on the fluid.
The temperature gradient $\gradient{T}$ in terms of the conservative variables can be found in \cref{chap:appendix-pde}.

@@

\sidetitle{Gravity}
We also support two different kind of optional source terms.
The first source term is gravity, given by
\begin{equation}\label{eq:source-gravity}
  \source[\Qj] = - g \bm{e_z} \cdot \Qj,
\end{equation}
where $\bm{e_z}$ is the unit vector pointing in $z$-direction~\cite{giraldo2008study}.

@@

\sidetitle{Chemical reactions}
The chemical reactive source term is
\begin{equation}\label{eq:source-reaction}
  \source[\QZ] = - K(T) \QZ.
\end{equation}
We use the simple reaction model
\begin{equation}\label{eq:reaction-model}
  K(T) = 
\begin{cases}
  - \frac{1}{\reactionTimescale} & T > \reactionTemperature\\
  0 & \text{otherwise}
\end{cases}  
\end{equation}
where $\reactionTimescale$ is the timescale of the reaction and $\reactionTemperature$ is the activation temperature~\cite{hidalgo2011ader,helzel2000modified}.

@@

The \aderdg{}-scheme is a predictor-corrector method.
We first compute a local time evolution of each cell in a predictor step and then add the influence of the neighbors in the corrector step.

\sidetitle{Reference cell}
For reasons of computational efficiency, we describe the scheme in terms of a reference cell.
We use the quad cell $\hat{\cell} = (0, 1)^\dimensions$ as space reference cell.
The mapping, which takes a point on the reference cell with coordinates $\hat{\bm{x}}$ and and returns its coordinates $\bm{x}$ in a grid cell $\cell$ is
\begin{equation}\label{eq:space-mapping}
  \bm{x} = \mapping (\hat{\bm{x}}) =
  \cellCenter(\cell) +
\begin{pmatrix}
\Delta x && \\
&\Delta y & \\
&&\Delta z&
\end{pmatrix}
  \begin{pmatrix}
    \hat{x} - 0.5\\
    \hat{y} - 0.5\\
    \hat{z} - 0.5
  \end{pmatrix}.
\end{equation}
The symbol $\mapping^{-1}$ denotes the inverse mapping.
It can be used to define a function $f(\bm{x})$ acting on a point $\bm{x}$ of a regular cell in terms of a function $\hat{f}(\bm{x})$ that acts on the corresponding point of the reference cell by the relation
\begin{equation}
  \label{eq:function-ref-cell}
  f(\bm{x}) = f\left( \mapping^{-1}(\bm{x}) \right) = \hat{f}(\hat{\bm{x}}).
\end{equation}

@@

The Jacobian determinant of $\mapping$ is simply the volume $V$ of the cell
\begin{equation}
  \label{eq:determinant-mapping}
  \volume = \operatorname{det}(\operatorname{Jacobian}\mapping) = \Delta x \, \Delta y \,\Delta z,
\end{equation}
the Jacobian determinant of $\timeMapping$ is $\Delta t$.
This is useful for evaluating derivatives by the chain rule and, similarly, integrals by substitution.
For example, we can evaluate a volume integral over a cell by
\begin{equation}
  \label{eq:integration-by-substitution}
  \intdcell{
f(\bm{x})
  }
  =
\volume \intdrefcell{
    \hat{f}(\hat{\bm{x}})
  }.
\end{equation}

\sidetitle{Basis functions}
The next step is the definition of basis functions for our cells.
We use Lagrange polynomials that are defined in the one-dimensional case for the reference interval $(0,1)$ by
\begin{align}
  \label{eq:lagrange-basis}
    \lagrangeRef(x) = \prod_{\substack{0 \leq j \leq N \\ i \neq j}} \frac{x - {\quadNode[j]}}{x_i - \quadNode[j]}.
\end{align}
We can then approximate functions as a sum over the support points $\quadNode[j]$
\begin{equation}
  \label{eq:langrange-expansion-1d}
  f(x) = \sum_{i = 0}^N f(\quadNode[i]) \lagrangeRef[i](x) = f(\quadNode[i]) \lagrangeRef[i](x),
\end{equation}
where we used the Einstein summation convention which implies summation over repeated free indices.
This convention is used from here on.
The support points $\quadNode[i]$ are chosen such that they coincide with the nodes of the Gauss-Legendre quadrature of order $N+1$ which we can use to evaluate integrals
\begin{equation}
  \label{eq:quadrature-1d}
  \int_0^1 f(\hat{x}) \dd{\hat{x}} = \sum_{i=0}^{N} \quadWeight[i] f(\quadNode[i]),
\end{equation}
where $\quadWeight[i]$ is the quadrature weight corresponding to the node $\quadNode[i]$.

The basis functions are similarly defined in $\dimensions$-dimensions by considering nodes that are the tensor-products of $d$ one-dimensional Gauss-Legendre quadrature nodes
\begin{equation}
  \label{eq:basis-space}
  \sbasisRef[][n] (\bm{x}) = \prod_{i=0}^{\dimensions} \lagrangeRef[n_i] (x_i).
\end{equation}
They are indexed by a multi-index $\bm{n} = (n_1, \ldots, n_\dimensions)$ that is linearized such that $n = (0, \ldots, \sbasisSize)$.
A $d$-dimensional function is then approximated by
\begin{equation}
  \label{eq:lagrange-expansion}
  f(\bm{x}) = \sum_{i = 0}^{\sbasisSize - 1} f(\quadNodeNd[i]) \lagrangeRef[i](\bm{x}) = f(\quadNodeNd[i]) \lagrangeRef[i](\bm{x}).
\end{equation}
Similarly to the one-dimensional case, we can evaluate integrals by
\begin{align}
  \label{eq:quadrature}
  \underbrace{\int_0^1 \cdots \int_0^1}_{\dimensions \text{ times}} f(\hat{\bm{x}}) \dd{\hat{\bm{x}}} &=
 \sum_{n=0}^{\sbasisSize-1} \quadWeight[n] f(\quadNodeNd[n]) \\
  \intertext{with}
  \quadWeight[n] &= \prod_{i}^{d} \quadWeight[n_i],
\end{align}
where the quadrature weights $\quadWeight[n]$ are the product of all relevant one-dimensional quadrature weights, using the aforementioned (linearized) multi-index $n$.

@@

\sidetitle{Cell basis}
We define the basis functions over a space-cell with an index $n = 0, \ldots, \sbasisSize - 1$ as
\begin{align}\label{eq:space-basis}
  \begin{split}
  \sbasis[][n] (\bm{x}) &= \sbasisRef[][n] \left( \mapping^{-1}(\bm{x}) \right).
  \end{split}
\end{align}
In addition to the space cells, we also introduce space-time cells $\cell \times [\curTime, \nextTime]$ with corresponding reference cell $(0,1)^\dimensions \times  [0,1]$.
A basis for these cells is given by
\begin{align}\label{eq:space-time-basis}
  \begin{split}
  \stbasisRef[][l,n](\hat{\bm{x}}, \hat{t}) &= \lagrange[l](t) \sbasis[][n] (x)\\
  \stbasis[][l,n] (\bm{x}, t) &= \stbasisRef \left( \mapping^{-1}(\bm{x}), \timeMapping^{-1}(t) \right)
  \end{split}
\end{align}
where the first index $l = 0, \ldots, N$ runs over the time.
Outside of elements, we define the basis functions to be equal to zero.

@@

\sidetitle{Predictor}
We are now ready to derive the predictor.
We first multiply the conservation law by a test function of the same function space as the basis, integrate over a space-time-element $\cell$ and replace all terms with their discrete approximations.
In particular, we replace the vector of conservative variables $\Q$ with the space-time predictor $\stpredictor[\cell]$ and arrive at
\begin{align}\label{eq:weak-pde-space-time}
\begin{split}
\intdt{\intdcell{
    \sttestfunction[\cell][v,l,n](\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
&+
\intdt{\intdcell{
    \sttestfunction[\cell][v,l,n](\bm{x}, t)
    \left(
      \divergence{\stflux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
\\&=
\intdt{\intdcell{
  \sttestfunction[\cell][v,l,n] \stsource(\bm{x}, t, \stpredictor)
}}.
\end{split}
\end{align}
%where the indices $v = (0, \ldots, \NVar-1)$, $n = (0, \ldots, \sbasisSize - 1)$ and $l = (0, \ldots, N)$ run over the variables, space and time.

Henceforth, we drop the cell index for all quantities and the spatial argument for the basis and test functions.
We integrate the first term of \cref{eq:weak-pde-space-time} by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the known discrete solution at time $t$.
This corresponds to upwinding in time~\cite{dumbser2008unified}.
Note that this neglects the interaction with neighboring cells.
The cell-local scheme is then
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \sttestfunction[][v,l,n] (\nextTime) \stpredictor[i+1](\nextTime)
}
&-
\intdt{\intdcell{
    \pdv{}{t} \sttestfunction[][v,l,n] (t) \stpredictor[i+1](t)
}}
= 
\intdcell{
  \sttestfunction[][v,l,n] (\curTime) \ssol (\curTime)
}
\\&+
\intdt{\intdcell{
    \sttestfunction[][v,l,n] (t) \divergence{\stflux(\stpredictor[i], \gradient{\stpredictor[i]})}
}}
\\&+
\intdt{\intdcell{
    \sttestfunction[][v,l,n] (t) \stsource(\stpredictor[i])
}}.\raisetag{0.8cm}
\end{split}
\end{align}
Here, we introduced the iteration counter $i$ as a subscript for the space-time predictor.
Using the expansions in space and time~\pcref{eq:cell-approx-space,eq:cell-approx-space-time} results in a local systems of equations that can be solved by a fixed-point iteration scheme.
We set the initial value for the space-time predictor as
\begin{equation}
  \label{eq:initial-guess}
\quad \stpredictorCoeff[0][v,l,n] = \ssolCoeff[][v,n]
\end{equation}
which corresponds to a stationary initial guess~\cite{dumbser2008unified}.
For an overview of possible initial guesses, see~\cite{dumbser2018efficient}.

@@

\sidetitle{Corrector}
The final step of the \aderdg{}-scheme is the corrector.
We multiply the system \cref{eq:conservation-law-gradient} by a space test function and integrate over the space-time control volume.
This results, yet again, in the weak formulation of the \pde{}
\begin{align}
  \begin{split}
  \label{eq:weak-pde}
\intdt{\intdcell{
\stestfunction[][v,n] \pdv{\Q}{t}
}}
&+
\intdt{\intdcell{
    \stestfunction[][v,n] \left( \div{\flux(\Q, \gradQ } \right)
}}
\\&=
\intdt{\intdcell{
    \stestfunction[][v,n] \source(\Q, \bm{x}, t)
}}.
  \end{split}
\end{align}

We now insert our discrete approximations and apply the space-basis expansion~\pcref{eq:cell-approx-space} to the discrete solution.
The first term can be integrated by parts in time.
This has the convenient result that only the boundary terms remain because the basis is constant in time.
We then integrate the flux divergence by parts in space and insert the Riemann solver to connect with the neighboring elements.
Finally, we need to solve
\newcommand{\massMatrixDef}{\intdcell{
  \stestfunction[][v,n] \sbasis[][v,n']
}}%
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\ssolCoeff[\nextTime][v,n'] - \ssolCoeff[\curTime][v,n'])
&+
\left(\intdt{\intdcellb{
      \stestfunction[][v,n] \Riemann(\stpredictor[-], \gradient{\stpredictor[-]}; \stpredictor[+], \gradient{\stpredictor[+]}) \cdot \normal
}}\right)
\\&-
\left(\intdt{\intdcell{
    \gradient{\stestfunction[][v,n]} \cdot  \stflux(\spredictor, \gradient{\spredictor})
}}\right)
\\&=
\left(\intdt{\intdcell{
      \stestfunction[][v,n] \stsource(\spredictor)
}}\right)
\end{split}\raisetag{0.8cm}
\end{align}
for the coefficients $\ssolCoeff[\nextTime][v,n]$ of the new solution.
Here the superscripts $\pm$ denote values that are extrapolated to the boundaries.
Again, the computational effort can be vastly reduced by collecting the integrals into matrices and precomputing as much as possible.
Using the orthogonality of the basis functions~\pcref{eq:basis-orthogonality} on the first term of \cref{eq:corrector}, it becomes clear that the correction is a simple, explicit computation.
% The first term is the mass tensor with entries
% \newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
% \begin{equation}
%   \label{eq:mass-matrix}
%   \massMatrix[v n n'] = \massMatrixDef = \volume \underbrace{\quadWeight[n] \delta_{n n'}}_{\text{diagonal}}
% \end{equation}
% which is, due to the orthogonal basis \pcref{eq:basis-orthogonality}, diagonal for each variable $v$ and thus trivial to invert.
% We can then insert the definition of our basis function into our equation and arrive at
% \begin{align}
%   \massMatrix \left( \ssolCoeff[t+1] - \ssolCoeff[t] \right) &=
%   \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
%   \intertext{with}
% a_{v,n} &= \sfluxCoeff[][n'] \intdcell{
%   \sbasis[][v,n'] \gradient{\sbasis[][v,n]}(t)
% }
% \\
% \label{eq:corrector-surface}
% b_{v,n} &= \intdcellb{
%     \stestfunction[][v,n] \Riemann \left( \spredictor[-], \gradient{\spredictor[-]}; \spredictor[+], \gradient{\spredictor[+]} \right) \cdot \normal
% } \\
% s_{v,n} &=
% \ssourceCoeff[][v,n'] \intdcell{
%   \sbasis[][v,n] \sbasis[][v,n']
% }
% ,
% \end{align}
% where we evaluated the time-integrals by \cref{eq:integration-by-substitution}.
% All quantities are now time-averaged~\pcref{eq:time-average} and are represented in the space basis~\pcref{eq:cell-approx-space}.
Note that we evaluate the Riemann problem for the time-averaged space-predictor in our implementation.
This is justified by the linearity of our numerical flux~\cite{dumbser2008unified,dumbser2010arbitrary}.
% The surface integral \pcref{eq:corrector-surface} can be evaluated by summing over the contributions of the faces of each cell, which are computed by integrating the extrapolated space predictor with Gaussian quadrature.
% Finally, we can solve \cref{eq:corrector} for the new unknows $\ssolCoeff[k+1]$
% \begin{align}\label{eq:update-predictor}
% \begin{split}
%   %\bm{u}^{n+1}
%   \ssolCoeff[k+1][v]
%   &= \ssolCoeff[k][v] + \Delta t (\operatorname{update}) \\
%   \operatorname{update} &= \massMatrix[v]^{-1} \left( \bm{a}_v - \bm{b}_v + \bm{s}_v \right)
% \end{split}
% \end{align}
% which is clearly of the form of a one-step scheme.

@@

\sidetitle{Outliers}
